---
title: "Data visualization"
output: html_document
date: '2022-08-17'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)

library(sf)

```

### Reach in date

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))

```


```{r}
ggplot(data = lizards) +
  geom_point(aes(x = total_length, y = weight))


```


### Histogram of total lengths for all lizards in the dataset

```{r}

ggplot(data = lizards, aes(x = total_length)) + 
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.5,
                 linetype = "dotted")

```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(
    shape = 22,
    color = "cyan4",
    fill = "yellow",
    size = 2,
    alpha = 0.4
  )
```

## Plotting based on variables with the aes()

```{r}

ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             alpha = 0.5) +
  theme_minimal()

```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point() +
  facet_wrap(~common_name, scales = "free")
```

### Reordering

# Find total count of lizards by name

```{r}

lizards_species_totals <- lizards %>% 
  group_by(common_name) %>% 
  summarise(count = n(),
            mean_length = mean(total_length))

lizards_species_totals <- lizards %>% 
  count(common_name)

# lizards__species_totals <- lizards %>% 
#   count(common_name, site, tail)

```

## Horizontal graphing
### Reordering!

```{r}

ggplot(data = lizards_species_totals, aes(y = fct_reorder(common_name, n),
                                          x = n)) +
  geom_col()


```

Create scatterplot, faceted by species, for total length vs weight. The color and shape of points should differ based on species (common_name). Change the theme. Make each faceted panel have flexible axes limits (scales should be free).


```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 shape = common_name)) +
  facet_wrap(~common_name, scales = "free") +
  theme_minimal() +
  labs(x = "Total length (in mm)",
       y = "Weight (in grams)",
       title = "Jornada Basin lizard sizes",
       subtitle = "All lizards follow the standard length-weight relationship, with western whiptails being largest",
       caption = "Data: Jornada Basin LTER, CCBY")

```

### Fun with scales

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple",
                       high = "orange")

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue", "green", "yellow"))

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black")

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",
                     mid = "white",
                     high = "orange",
                     midpoint = 150,
                     breaks = c(50, 75, 150, 180, 220, 280))


ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, to = 400, by = 80))

```

# Paletteer

Make a horizontal box plot with common name on y-axis, total length on x-axis, with boxplot fill changing based on the common name

```{r}
lizards_median <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

# Creating a new column which overwrites existing column, recast common name as a factor not char, then reorders them according to the median of total length for each of the different levels of species by common_name (factor levels)

ggplot(data = lizards, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name))

ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::default_gsea") +
  theme_minimal()

ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "palettetown::pikachu") +
  theme_minimal()

```

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.x = element_line(color = "gold"),
        axis.title.x = element_text(color = "peachpuff"),
        axis.text.x = element_text(color = "cyan4"),
        text = element_text(size = 18),
        panel.background = element_rect(color = "gold",
                                        fill = "peachpuff")) +
  annotate("text", x = 300, y = 50, label = "Woooooo", color = "blue") +
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)

```


```{r}

wws_lizards <- lizards %>% 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = wws_lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num))

gapminder %>% 
  filter(year == 2002, continent == "Europe") %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country, size = 2.5), show.legend = FALSE)

```

```{r}

ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  gghighlight(toe_num == 250, label_key = toe_num)

```


```{r}
# sf stands for simple features, which is a way to store spacial data where spatial data sticks to attributes

jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() +
  geom_sf(data = jornada_veg,
          aes(fill = Name)) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")


  
```














